HERMES : Hybrid ERror-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2021-12-01 hermes_paper/hermes_paper.tex
a few updates

Following this work, we present in this paper HERMES, a new hybrid recurrent model time series forecasting with inclusion of external weak signals. By adding our fashion external weak signals in our proposed framework, we show the real potential of the hybrid approach: a global neural network, able to leverage large amount of data, deal with any kind of external weak signals, learn context and finally correct weaknesses and errors of any parametric models.
%Our new hybrid framework is divided in two parts. As a first step, we compute a per-time-series forecast for each trend using any kind of parametric models. Then, gathering information of the first prediction and external variables, a global error corrector neural network is trained to correct weaknesses of the first forecast.

The paper is organized as follow. Section~\ref{sec:hybrid} describes our hybrid framework which consists of a per-time-series parametric model and a recurrent neural network combining the target sequences with external weak signals. Then, we present our new fashion dataset in Section~\ref{sec:dataset}. Section~\ref{sec:exp} describes our results and comparisons with several benchmarks. Finally, we conclude our paper and detail research perspectives in Section~\ref{sec:discussion}.

\section{Hybrid model with external signals}
\label{sec:hybrid}
%We introduce a new hybrid approach for time series forecasting  composed of two parts: a first per time series parametric model with seasonal component and a global RNN. We use a per time series statistical model to learn local behaviour and to normalize sequences before feeding them to the neural network.
%We introduce a new hybrid approach for time series forecasting  composed of two parts: a first per time series parametric model with seasonal component and a global neural network framework composed of two RNN. We use a per time series statistical model to learn local behaviour, to normalize sequences by removing trends and seasonality,  and to compute a first forecast. Then, we use two recurrent neural networks and weak signals to correct, if needed, the predictions provided by the first collection of parametric models.
We introduce a new hybrid approach for time series forecasting  composed of two parts: a first per time series parametric model and a global error-corrector neural network train on all the time series. We use a per time series parametric model to learn local behaviour, to normalize sequences by removing trends and seasonality,  and to compute a first forecast. Then, gathering information of the first prediction and external variables, we use a recurrent neural networks to correct the predictions provided by the first collection of per-time-series models.

@@

\subsection{Per time series predictor}
%The first step in our additive hybrid model is to consider, for each time series, an exponential smoothing  statistical model.
%The time-series-specific models compute, for each sequence, two components,  level  and  seasonality. For all $1\leqslant n \leqslant N$ and $0\leqslant t \leqslant T$, let $\lvl_t^n$ be the value of the $n$-th level component at time $t$ and  $\fulllvl^n = \{\lvl_t^n\}_{0\leqslant t \leqslant T}$ be all the values of this level component. For all $1\leqslant n \leqslant N$ and $-m+1\leqslant t \leqslant T$, let $\seas_t^n$ be the value of the $n$-th seasonality component at time $t$ and  $\fullseas^n = \{\seas_t^n\}_{-m+1\leqslant t \leqslant T}$ be all the values of this season component. $m \in \mathbb{N}$ is the seasonality (4 for quarterly, 12 for monthly or 52 for weekly time series).
The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n(.;statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
$$

@@

Let $\window \in \mathbb{N}$ be the window size, usually this window is equal to the forecast horizon $\window = \lag$. We define as our RNN input the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnninput^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$, where, for all $1\leqslant n \leqslant N$ and $1\leqslant i \leqslant w$, 
$$
\rnninput^{n,T}_{T-w+i|T} = \frac{\ts^n_{T-w+i} -\tspred^{pred,n}_{T+i|T}}{\meants^n_T}\,,
$$
where $ \meants^n_T = \sum_{i = 1}^{w}  \ts^n_{T-w+i}/w$.

@@

Using well-fitted time-series-specific parametric models, our hybrid network corrects the first form of weakness and achieves striking accuracy on our dataset. Then, for the second form of weakness, the shared recurrent neural network relies on the strength of machine learning methods at leveraging a huge quantities of heterogeneous data.
In addition to the $N$ target time series, we consider $K \times N$ external sequences indexed from $0$ to $T$. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ linked to the sequence $\fullts^n$. We note $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ all the values of the family of weak signals linked to the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the family. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, we define a new input for our two RNN,   
$$
\fullconcatinput^n_T = \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
$$
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ our final model becomes:
%$$
%\tspred^n_{T+i|T} 
%= \etspred^{ets,n}_{T+i|T} +  \rnnmodel(\fullconcatinput^n_T)_{T+i} \times \meants^n_T
%$$
\begin{equation}
\label{eq:withws:full:model}
\tspred^n_{T+i|T}  = \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\fullconcatinput^n_T;\rnnparam)_i \cdot \meants^n_T\
\end{equation}

@@

To train our hybrid model, we split our dataset into three blocks, train, eval and test sets. We use the 3 first years as our train set, we keep the 4th year for our eval set and we hide the last year for our test set. We train our hyrid model to compute a one-year ahead prediction with $\window$ equal to 104 and a $\lag$ equal to 52.
Using the two first year of our train set, we fit a first per-time-series parametric model for each time series. With the resulting collection of local models, we compute a first forecast of the third year for each sequence. Using these forecasts, we preprocess our time series and finally determine our RNN inputs. We can, then, train our RNN at understanding and correcting the previous collection of parametric model's forecasts. In the same way, for the eval set, we fit a second time our per-time-series predictors using the three first years and we forecast the fourth year. We preprocess our time series and we evaluate how our error-corrector RNN is able to improve the first collection of predictions of the fourth year.
At the end, we repeat the same process for the last year and compute our final accuracy measures. An example of our splitting is shown in  Figure~\ref{fig:train_eval_test_set}.
%We train our hyrid model to compute a one-year ahead prediction with $\window$ equal to 104 and a $\lag$ equal to 52. For each time series, we fitted 3 per-time-series parametric models: We fit the first one using the 2 first years and compute a forecast of the thrid year.

@@

For the first parametric per-time-series models, we used existing Python or R librarys to estimate the different parameters $\statparam^n$. There are then fixed during the training of the recurrent models. The proposed hybrid model introduced in this numerical section is the TBATS model of \cite{doi:10.1198/jasa.2011.tm09771}. This model is also compared to an hybrid model with an additive exponential smoothing model as a reference close to \cite{smyl2020hybrid}. For the neural network part, we built an architecture summarized in Figure~\ref{fig:rnnarchi}. It is composed of 3 LSTM layer of shape 50 and a final Dense layer to provide the correct output dimension. We use a classical Adam otpimizer with a learning rate equal to 0.005 and a batch size equal to 64. For our loss function, we tested several metrics and we finally converged on the following one :

@@

\begin{figure}
  \centering
    \includegraphics[width=1.\linewidth]{figure/train_eval_test_set}
  \caption{$\classifier$ and $\predictor$ architecture.}
\label{fig:train_eval_test_set}
\end{figure}

@@

As benchmarks, we chose several well-known statistical methods and deep learning approaches. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels, tbats} , we compute for each time series a prediction with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. \textit{snaive} forecasts are only the repetition of the last same past period. \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. \textit{stlm} approach uses a multiplicative decomposition and models the seasonally adjusted time series with an exponential smoothing model. \textit{Thetam} model decomposes the original signal in $\theta$-lines, predicts each one separately and recomposes them to produce the final forecast. \textit{tbats} is a powerful statistical model using a trigonometrical seasonality modelization. The last one \textit{auto.arima} is the R implementation of the ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \citep{hyndman2020package}. As a deep learning approach, we considered a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52.

Against these models, we propose two versions of our hybrid model (hermes) called respectively hermes-ets and herles-tbats according to the per-time-series model choice.
%Against these models, we propose two versions of our regime switching hybrid recurrent model \textit{rshm}. The first one, that we name \textit{rshm-ets}, is the hybrid model with an additive exponential smoothing model as parametric model. For the second one, we use the TBATS approach as paramteric model and we named it \textit{rshm-tbats}. 
%To reach a higher accuracy level we also propose an ensembling (\textit{ensembling} of our hybrid approach using the FFORMA method created during the M4 competition, see \citep{fforma}). 
For our two versions, we propose a version without our weak signal called \textit{hermes-ets} and \textit{hermes-tbats}. Then we evaluate our hybrid framework with the inclusion of our weak signals (ws) referred to as \textit{rshm-ets-ws} and \textit{rshm-tbats-ws}. In order to provide a fair comparison, we also train a \textit{lstm} with the weak signals named \textit{lstm-ws}.

>>>>

Following this work, we present in this paper HERMES, a new hybrid recurrent model for time series forecasting with inclusion of external signals. Our new framework is divided in two levels: local predictors and a global corrector.  In a first time, given a group of sequence, we compute a per-time-series prediction using any kind of parametric models. In a second time, we train a global neural network to understand, evaluate and correct the forecast weaknesses of the first collection of models. By adding external weak signals in our proposed framework, we show the real potential of the hybrid approach: a global neural network, able to leverage large amount of data, deal with any kind of external weak signals, learn context and finally correct weaknesses and errors of any parametric models.
%Our new hybrid framework is divided in two parts. As a first step, we compute a per-time-series forecast for each trend using any kind of parametric models. Then, gathering information of the first prediction and external variables, a global error corrector neural network is trained to correct weaknesses of the first forecast.

%Section~\ref{sec:hybrid} describes our hybrid framework which consists of a per-time-series parametric model and a recurrent neural network combining the target sequences with external weak signals

The paper is organized as follow. Section~\ref{sec:hybrid} describes our hybrid framework. Then, we present our new fashion dataset in Section~\ref{sec:dataset}. Section~\ref{sec:exp} describes our results and comparisons with several benchmarks. Finally, we conclude our paper and detail research perspectives in Section~\ref{sec:discussion} and Section~\ref{sec:conclusion}.

\section{Hybrid model with external signals}
\label{sec:hybrid}
%We introduce a new hybrid approach for time series forecasting  composed of two parts: a first per time series parametric model with seasonal component and a global RNN. We use a per time series statistical model to learn local behaviour and to normalize sequences before feeding them to the neural network.
%We introduce a new hybrid approach for time series forecasting  composed of two parts: a first per time series parametric model with seasonal component and a global neural network framework composed of two RNN. We use a per time series statistical model to learn local behaviour, to normalize sequences by removing trends and seasonality,  and to compute a first forecast. Then, we use two recurrent neural networks and weak signals to correct, if needed, the predictions provided by the first collection of parametric models.
We introduce a new hybrid approach for time series forecasting  composed of two parts: first per-time-series parametric models, the predictors, and a global error-corrector neural network train on all the time series, the corrector. We use  per-time-series parametric models to learn local behaviour, to normalize sequences by removing trends and seasonality,  and to compute a first forecast. Then, gathering information of the first predictions and external variables, we use a recurrent neural networks to correct the predictions provided by the first collection of per-time-series models.

@@

\subsection{Per-time-series predictors}
%The first step in our additive hybrid model is to consider, for each time series, an exponential smoothing  statistical model.
%The time-series-specific models compute, for each sequence, two components,  level  and  seasonality. For all $1\leqslant n \leqslant N$ and $0\leqslant t \leqslant T$, let $\lvl_t^n$ be the value of the $n$-th level component at time $t$ and  $\fulllvl^n = \{\lvl_t^n\}_{0\leqslant t \leqslant T}$ be all the values of this level component. For all $1\leqslant n \leqslant N$ and $-m+1\leqslant t \leqslant T$, let $\seas_t^n$ be the value of the $n$-th seasonality component at time $t$ and  $\fullseas^n = \{\seas_t^n\}_{-m+1\leqslant t \leqslant T}$ be all the values of this season component. $m \in \mathbb{N}$ is the seasonality (4 for quarterly, 12 for monthly or 52 for weekly time series).
The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n(.;\statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
$$

@@

Let $\window \in \mathbb{N}$ be the window size, usually this window is proportional to the forecast horizon $\window \propto \lag$. We define as our RNN input the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnninput^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$, where, for all $1\leqslant n \leqslant N$ and $1\leqslant i \leqslant w$, 
$$
\rnninput^{n,T}_{T-w+i|T} = \frac{\ts^n_{T-w+i} -\tspred^{pred,n}_{T+i|T}}{\meants^n_T}\,,
$$
where $ \meants^n_T = \frac{1}{w}\sum_{i = 1}^{w}\ts^n_{T-w+i}$.

@@

Using well-fitted time-series-specific parametric models, our hybrid network corrects the first form of weakness and achieves striking accuracy on our dataset. Then, for the second form of weakness, the shared recurrent neural network relies on the strength of machine learning methods at leveraging a huge quantities of heterogeneous data.
In addition to the $N$ target time series, we consider $K \times N$ external sequences indexed from $0$ to $T$. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ linked to the sequence $\fullts^n$. We note $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ all the values of the family of weak signals linked to the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the family. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, we define a new input for our RNN,   
$$
\fullconcatinput^n_T = \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
$$
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ our final model becomes:
%$$
%\tspred^n_{T+i|T} 
%= \etspred^{ets,n}_{T+i|T} +  \rnnmodel(\fullconcatinput^n_T)_{T+i} \times \meants^n_T
%$$
\begin{equation}
\label{eq:withws:full:model}
\tspred^n_{T+i|T}  = \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\fullconcatinput^n_T;\rnnparam)_i \cdot \meants^n_T\
\end{equation}

@@

To train our hybrid model, we split our dataset into three blocks, train, eval and test sets. We use the 3 first years as our train set, we keep the 4th year for our eval set and we hide the last year for our test set. We train our hyrid model to compute a one-year ahead prediction, $\lag$ equal to 52, and we set the window size $\window$ equal to 104.
Using the two first year of our train set, we fit a first per-time-series parametric model for each time series. With the resulting collection of local models, we compute a first forecast of the third year for each sequence. Using these forecasts, we preprocess our time series and finally determine our RNN inputs. We can, then, train our RNN at understanding and correcting the previous collection of parametric model's forecasts. In the same way, for the eval set, we fit a second time our per-time-series predictors using the three first years and we forecast the fourth year. We preprocess our time series and we evaluate how our error-corrector RNN is able to improve the first collection of predictions of the fourth year.
At the end, we repeat the same process for the last year and compute our final accuracy measures. An example of our splitting is shown in  Figure~\ref{fig:train_eval_test_set}.
%We train our hyrid model to compute a one-year ahead prediction with $\window$ equal to 104 and a $\lag$ equal to 52. For each time series, we fitted 3 per-time-series parametric models: We fit the first one using the 2 first years and compute a forecast of the thrid year.

@@

For the first parametric per-time-series models, we used existing Python or R librarys to estimate the different parameters $\statparam^n$. These parameters could move during the training of the hybrid model proposed during the M4 competition. In our framework, there are fixed during the training of the recurrent models. In this section we propose two versions of HERMES depending of the choice of local parametric models. The first one use as predictors an additive exponential smoothing model as a reference close to \cite{smyl2020hybrid}. The second one use as predictors the TBATS model of \cite{doi:10.1198/jasa.2011.tm09771} and  achieves striking accuracy results on our fashion dataset. For the neural network part, we built an architecture summarized in Figure~\ref{fig:rnn_architecture}. It is composed of 3 LSTM layer of shape 50 and a final Dense layer to provide the correct output dimension. We use a classical Adam optimizer with a learning rate equal to 0.005 and a batch size equal to 64. For our loss function, we tested several metrics and we finally converged on the following one :

@@

\begin{figure}
  \centering
    \includegraphics[width=1.\linewidth]{figure/train_eval_test_set}
  \caption{Temporal split for our training process}
\label{fig:train_eval_test_set}
\end{figure}

@@

As benchmarks, we chose several well-known statistical methods and deep learning approaches. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels},  \texttt{tbats} , we compute for each time series a prediction with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. The forecast of the \textit{snaive} method is only the repetition of the last same past period. \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. \textit{stlm} approach uses a multiplicative decomposition and models the seasonally adjusted time series with an exponential smoothing model. \textit{Thetam} model decomposes the original signal in $\theta$-lines, predicts each one separately and recomposes them to produce the final forecast. \textit{tbats} is a powerful statistical model using a trigonometrical seasonality modelization. The last one \textit{auto.arima} is the R implementation of the ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \citep{hyndman2020package}. As a deep learning approach, we considered a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52.

Against these models, we propose two versions of our hybrid model (hermes) called respectively hermes-ets and herles-tbats according to the per-time-series model choice.
%Against these models, we propose two versions of our regime switching hybrid recurrent model \textit{rshm}. The first one, that we name \textit{rshm-ets}, is the hybrid model with an additive exponential smoothing model as parametric model. For the second one, we use the TBATS approach as paramteric model and we named it \textit{rshm-tbats}. 
%To reach a higher accuracy level we also propose an ensembling (\textit{ensembling} of our hybrid approach using the FFORMA method created during the M4 competition, see \citep{fforma}). 
For our two versions, we propose a version without our weak signal called \textit{hermes-ets} and \textit{hermes-tbats}. Then we evaluate our hybrid framework with the inclusion of our weak signals (ws) referred to as \textit{hermes-ets-ws} and \textit{hermes-tbats-ws}. In order to provide a fair comparison, we also train a \textit{lstm} with the weak signals named \textit{lstm-ws}.
